---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First of all load the dataset and convert the data into convenient R data types.
```{r, echo=TRUE}
# load dataset
activity.df <- read.csv("activity.csv"
                            ,header=T
                            ,sep=","
                            ,na.string="NA"
                            ,stringsAsFactors=F)

# convert Date and numeric fields in Date/Integer R classes
activity.df$steps <- as.integer(activity.df$steps)
activity.df$date <- as.Date(activity.df$date,"%Y-%m-%d")
activity.df$interval <- as.integer(activity.df$interval)

```
Next step, remove the NA rows from the initial dataset to perform a fist cut analysis

```{r, echo=TRUE}
# remove NAs from first cut data analysis
activity.nona <- activity.df[complete.cases(activity.df),]

```
Next plot a histogram of the total number of steps taken each day
and calculate and report the mean and median total number of steps taken per day

```{r, echo=TRUE}
# tot steps by date
totsteps.by.date<-aggregate(activity.nona$steps, by=list(activity.nona$date), FUN=sum)
names(totsteps.by.date)<-c("date","tot.steps")
# plot total steps by date

# open graphic device
png(filename = "totstepsbydate.png", width = 480, height = 480, units = "px")

# plot bar chart
barplot(totsteps.by.date$tot.steps,names=totsteps.by.date$date,col="orangered3",xlab="Date",ylab="Tot. Steps")
# set chart title
title("Total number of steps by date")

# median steps per day
median.steps<-median(totsteps.by.date$tot.steps)
abline(h = median.steps, col = "blue", lwd = 2)
# mean steps per day
mean.steps<-mean(totsteps.by.date$tot.steps)
abline(h = mean.steps, col = "green", lwd = 2)

legend("topleft",lty = c(1,1,1),col= c("blue","green"), legend=c("median","mean"))
#close graphic device
dev.off()

```
## What is mean total number of steps taken per day?
The mean and median total number of steps taken per day
are 10766 and 10765 respectively.

Next, investigate what is the average daily activity pattern
by analyzing the 5-minute interval and the average number of steps taken, averaged across all days
```{r, echo=TRUE}
# avg steps by interval
avgsteps.by.interval<-aggregate(activity.nona$steps, by=list(activity.nona$interval), FUN=mean)
names(avgsteps.by.interval)<-c("interval","avg.steps")

# open graphic device
png(filename = "avgstepsbyinterval.png", width = 480, height = 480, units = "px")

#plot chart
with(avgsteps.by.interval,plot(interval,avg.steps
                          ,type="l"
                          ,xlab="5-minutes interval"
                          ,ylab="Avg number of steps"))
# set chart title
title("Average number of steps by 5-minutes interval")
#close graphic device
dev.off()

```

## What is the average daily activity pattern?
Next, report that the 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps is

```{r, echo=TRUE}

# identify the interval with max number of steps (on average)
top.interval<-avgsteps.by.interval[which.max(avgsteps.by.interval[,2]),]

```
interval avg.steps
 8:35     206.1698

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
